---
title: "Pumba datamining"
author: "Roman Mylonas"
date: "2 Oct 2019"
output: html_document
---

```{r message=FALSE, echo=FALSE, warning=FALSE}
suppressPackageStartupMessages(library("ggplot2"))

# load results
res <- path <- ("/Users/rmylonas/tmp/datamining_pumba/results/homodimers_1569571862.15343.RData")
load(res <- path)

sample <- "HCT"
res <- results[[sample]]

show_mass_range <- c(10, 300)

peak_selection <- "closest_peak_masses" #"highest_peak_masses"
peak_selection_term <- "CLOSEST" # "HIGHEST"
```

# Data processing
- Data is extracted from Pumba
- Peaks are detected
- Data is annotated from additional sources (e.g. UniProt, SCOP)

# Peak detection
- Peaks below an absolute normalised intensity of 1e-08 are ignored.
- Only peaks >= 10% compared to strongest peak are retained.
- In between adjacent peaks the curve has to go through <= 80% of last peak, otherwise only the strongest peak is retained.  

# Plot theoretical vs observed masses
The observed masses correspond to the *`r peak_selection_term`* peak.

### Protein intesities

```{r message=FALSE, echo=FALSE, warning=FALSE}
ggplot(res[order(res$prot_intensities, decreasing=FALSE),], aes_string(x="theo_weights", y=peak_selection, color="prot_intensities")) +
  geom_point() +
  scale_colour_gradient(low = "white", high = "#08519c", trans="log") +
  scale_x_log10() +
  scale_y_log10() +
  coord_cartesian(xlim = show_mass_range, ylim = show_mass_range) +
  geom_abline(intercept = 0, slope = 1) +
  theme_bw()
```

## Sequence properties
Sequence properties are computed using the R package *Peptides*. 

### Isoelectric point (pI)
- pos > 9
- neg < 5
- neutral are between 5 and 9

```{r message=FALSE, echo=FALSE, warning=FALSE}
res$pI_status <- "neutral"
res$pI_status[res$pIs > 9] <- "pos"
res$pI_status[res$pIs < 5] <- "neg"

ggplot(res, aes_string(x="theo_weights", y=peak_selection, color="pI_status")) +
  geom_point(alpha=0.05) +
  geom_point(data=res[res$pI_status == "neutral", ]) +
  scale_x_log10() +
  scale_y_log10() +
  coord_cartesian(xlim=show_mass_range,ylim=show_mass_range) +
  geom_abline(intercept=0, slope=1) +
  theme_bw()

ggplot(res, aes_string(x="theo_weights", y=peak_selection, color="pI_status")) +
  geom_point(alpha=0.05) +
  geom_point(data=res[res$pI_status == "pos", ]) +
  scale_x_log10() +
  scale_y_log10() +
  coord_cartesian(xlim=show_mass_range,ylim=show_mass_range) +
  geom_abline(intercept=0, slope=1) +
  theme_bw()

ggplot(res, aes_string(x="theo_weights", y=peak_selection, color="pI_status")) +
  geom_point(alpha=0.05) +
  geom_point(data=res[res$pI_status == "neg", ]) +
  scale_x_log10() +
  scale_y_log10() +
  coord_cartesian(xlim=show_mass_range,ylim=show_mass_range) +
  geom_abline(intercept=0, slope=1) +
  theme_bw()
```

### Hydrophobicity by Hopp Woods
- pos > 0.3
- neg < -0.3 
- neutral are between -0.3 and 0.3

```{r message=FALSE, echo=FALSE, warning=FALSE}
res$hydrophobicity <- "neutral"
res$hydrophobicity[res$hydrophobicities > 0.3] <- "pos"
res$hydrophobicity[res$hydrophobicities < -0.3] <- "neg"

ggplot(res, aes_string(x="theo_weights", y=peak_selection, color="hydrophobicity")) +
  geom_point(alpha=0.05) +
  geom_point(data=res[res$hydrophobicity == "neutral", ]) +
  scale_x_log10() +
  scale_y_log10() +
  coord_cartesian(xlim=show_mass_range,ylim=show_mass_range) +
  geom_abline(intercept=0, slope=1) +
  theme_bw()

ggplot(res, aes_string(x="theo_weights", y=peak_selection, color="hydrophobicity")) +
  geom_point(alpha=0.05) +
  geom_point(data=res[res$hydrophobicity == "pos", ]) +
  scale_x_log10() +
  scale_y_log10() +
  coord_cartesian(xlim=show_mass_range,ylim=show_mass_range) +
  geom_abline(intercept=0, slope=1) +
  theme_bw()

ggplot(res, aes_string(x="theo_weights", y=peak_selection, color="hydrophobicity")) +
  geom_point(alpha=0.05) +
  geom_point(data=res[res$hydrophobicity == "neg", ]) +
  scale_x_log10() +
  scale_y_log10() +
  coord_cartesian(xlim=show_mass_range,ylim=show_mass_range) +
  geom_abline(intercept=0, slope=1) +
  theme_bw()
```

## Sequence properties
Parse UniProt annotations from various fields (e.g. locations).

### Mitochondrial proteins

```{r message=FALSE, echo=FALSE, warning=FALSE}
# location term
location_term <- "mitochond"
  
neg_location_term <- paste0("not ", location_term)
res$color <- neg_location_term
res$color[grep(location_term, res$location, ignore.case = TRUE)] <- location_term

# colors
col_vec <- c("grey", "red")
names(col_vec) <- c(neg_location_term, location_term)

p <- ggplot(data=res[order(res$color),], aes_string(x="theo_weights", y=peak_selection, colour="color")) 
p <- p + geom_point(data = res[res$color == neg_location_term,], alpha=0.2)
p <- p + geom_point(data = res[res$color == location_term,], alpha=0.7)
p <- p + scale_x_log10()
p <- p + scale_y_log10()
p <- p + coord_cartesian(xlim=show_mass_range,ylim=show_mass_range)
p <- p + geom_abline(intercept=0, slope=1)
p <- p + scale_colour_manual(values=col_vec)
p <- p + theme_bw()
p <- p + ggtitle(sample)
print(p)
```

### Nucleus proteins

```{r message=FALSE, echo=FALSE, warning=FALSE}
# location term
location_term <- "nucleus"
  
neg_location_term <- paste0("not ", location_term)
res$color <- neg_location_term
res$color[grep(location_term, res$location, ignore.case = TRUE)] <- location_term

# colors
col_vec <- c("grey", "red")
names(col_vec) <- c(neg_location_term, location_term)

p <- ggplot(data=res[order(res$color),], aes_string(x="theo_weights", y=peak_selection, colour="color")) 
p <- p + geom_point(data = res[res$color == neg_location_term,], alpha=0.2)
p <- p + geom_point(data = res[res$color == location_term,], alpha=0.7)
p <- p + scale_x_log10()
p <- p + scale_y_log10()
p <- p + coord_cartesian(xlim=show_mass_range,ylim=show_mass_range)
p <- p + geom_abline(intercept=0, slope=1)
p <- p + scale_colour_manual(values=col_vec)
p <- p + theme_bw()
p <- p + ggtitle(sample)
print(p)
```

### Glycosylated proteins

```{r message=FALSE, echo=FALSE, warning=FALSE}
neg_location_term <- "no glycosylation"
location_term <- "glycosylated"

res$color <- neg_location_term
res$color[res$glycosylations != ""] <- location_term

# colors
col_vec <- c("grey", "red")
names(col_vec) <- c(neg_location_term, location_term)

p <- ggplot(data=res[order(res$color),], aes_string(x="theo_weights", y=peak_selection, colour="color")) 
p <- p + geom_point(data = res[res$color == neg_location_term,], alpha=0.2)
p <- p + geom_point(data = res[res$color == location_term,], alpha=0.7)
p <- p + scale_x_log10()
p <- p + scale_y_log10()
p <- p + coord_cartesian(xlim=show_mass_range,ylim=show_mass_range)
p <- p + geom_abline(intercept=0, slope=1)
p <- p + scale_colour_manual(values=col_vec)
p <- p + theme_bw()
p <- p + ggtitle(sample)
print(p)
```


### Ubiquitinated proteins (glycyl lysine)

```{r message=FALSE, echo=FALSE, warning=FALSE}
ptm <- "glycyl"
neg_label <- paste0("no ", ptm)
pos_label <- ptm

res$color <- neg_label
res$color[grep(ptm, res$ptms, ignore.case = TRUE)] <- pos_label

# colors
col_vec <- c("grey", "red")
names(col_vec) <- c(neg_label, pos_label)

p <- ggplot(data=res[order(res$color),], aes_string(x="theo_weights", y=peak_selection, colour="color")) 
p <- p + geom_point(data = res[res$color == neg_label,], alpha=0.2)
p <- p + geom_point(data = res[res$color != neg_label,], alpha=0.7)
p <- p + scale_x_log10()
p <- p + scale_y_log10()
p <- p + coord_cartesian(xlim=show_mass_range,ylim=show_mass_range)
p <- p + geom_abline(intercept=0, slope=1)
p <- p + scale_colour_manual(values=col_vec)
p <- p + theme_bw()
p <- p + ggtitle(sample)
print(p)
```

### Signal peptides

```{r message=FALSE, echo=FALSE, warning=FALSE}

neg_label <- "no signal peptide"
pos_label <- "signal peptide"

res$color <- neg_label
res$color[res$signal_peps != ""] <- pos_label

# colors
col_vec <- c("grey", "red")
names(col_vec) <- c(neg_label, pos_label)

p <- ggplot(data=res[order(res$color),], aes_string(x="theo_weights", y=peak_selection, colour="color")) 
p <- p + geom_point(data = res[res$color == neg_label,], alpha=0.2)
p <- p + geom_point(data = res[res$color != neg_label,], alpha=0.7)
p <- p + scale_x_log10()
p <- p + scale_y_log10()
p <- p + coord_cartesian(xlim=show_mass_range,ylim=show_mass_range)
p <- p + geom_abline(intercept=0, slope=1)
p <- p + scale_colour_manual(values=col_vec)
p <- p + theme_bw()
p <- p + ggtitle(sample)
print(p)
```  
  
### Phospho

```{r message=FALSE, echo=FALSE, warning=FALSE}
ptm <- "phospho"
neg_label <- paste0("no ", ptm)
pos_label <- ptm

res$color <- neg_label
res$color[grep(ptm, res$ptms, ignore.case = TRUE)] <- pos_label

# colors
col_vec <- c("grey", "red")
names(col_vec) <- c(neg_label, pos_label)

p <- ggplot(data=res[order(res$color),], aes_string(x="theo_weights", y=peak_selection, colour="color")) 
p <- p + geom_point(data = res[res$color == neg_label,], alpha=0.2)
p <- p + geom_point(data = res[res$color != neg_label,], alpha=0.7)
p <- p + scale_x_log10()
p <- p + scale_y_log10()
p <- p + coord_cartesian(xlim=show_mass_range,ylim=show_mass_range)
p <- p + geom_abline(intercept=0, slope=1)
p <- p + scale_colour_manual(values=col_vec)
p <- p + theme_bw()
p <- p + ggtitle(sample)
print(p)
```
  
  
## SCOP (Structural Classification of Proteins)
- Use approach described in *"Girstmair, Hannah, et al. "The Hsp90 isoforms from S. cerevisiae differ in structure, function and client range." Nature communications 10.1 (2019): 1-15."*.
- Search sequences against SCOP databases using blast-scop (v. 2.0.3).
- Keep entries with E-value <= 0.001.

```{r message=FALSE, echo=FALSE, warning=FALSE}
scop_classes <- read.csv("/Users/rmylonas/Work/PAF/projects/pumba/data/datamining/scop/SCOP_classes.csv")

for(scop_class_short in scop_classes$class){
  scop_class_name <- as.character(scop_classes$name[scop_classes$class == scop_class_short])
  
  # Membrane class from SCOP
  res$is_scop_class <- unlist(lapply(res$scop_classes, function(x){ if(length(grep(scop_class_short, x)) > 0) scop_class_name else ""}))
  
  show_mass_range <- c(10, 300)
  
  p <- ggplot(res[order(res$is_scop_class),], aes_string(x="theo_weights", y=peak_selection, color="is_scop_class")) +
    geom_point(data=res[res$is_scop_class == "", ], alpha=0.1, color="grey") +
    geom_point(data=res[res$is_scop_class != "", ]) +
    scale_x_log10() +
    scale_y_log10() +
    coord_cartesian(xlim=show_mass_range,ylim=show_mass_range) +
    geom_abline(intercept=0, slope=1) +
    ggtitle(scop_class_name) +
    theme_bw()
  
  print(p)
}
```

## Fragmentation

### Slices with protein identification

Plot in how many slices the a protein was found (as percentage).

#### Logarithmic scale

```{r message=FALSE, echo=FALSE, warning=FALSE}
ggplot(res[order(res$perc_slices, decreasing=FALSE),], aes_string(x="theo_weights", y=peak_selection, color="perc_slices")) +
  geom_point() +
  scale_colour_gradient(low = "white", high = "black", trans="log") +
  scale_x_log10() +
  scale_y_log10() +
  coord_cartesian(xlim = show_mass_range, ylim = show_mass_range) +
  geom_abline(intercept = 0, slope = 1) +
  theme_bw()

```

#### Linear scale

```{r message=FALSE, echo=FALSE, warning=FALSE}

ggplot(res[order(res$perc_slices, decreasing=FALSE),], aes_string(x="theo_weights", y=peak_selection, color="perc_slices")) +
  geom_point() +
  scale_colour_gradient(low = "white", high = "black") +
  scale_x_log10() +
  scale_y_log10() +
  coord_cartesian(xlim = show_mass_range, ylim = show_mass_range) +
  geom_abline(intercept = 0, slope = 1) +
  theme_bw()
```

### Number of peaks

Plot the number of identified peaks per protein.

```{r message=FALSE, echo=FALSE, warning=FALSE}
ggplot(res[order(res$nr_peaks, decreasing=FALSE),], aes_string(x="theo_weights", y=peak_selection, color="nr_peaks")) +
  geom_point() +
  scale_colour_gradient(low = "#bdd7e7", high = "#08519c") +
  scale_x_log10() +
  scale_y_log10() +
  coord_cartesian(xlim = show_mass_range, ylim = show_mass_range) +
  geom_abline(intercept = 0, slope = 1) +
  theme_bw()
```
  


