---
title: "Proteins which elute differently then expecteed"
author: "RM"
date: "11/24/2019"
output: html_document
---

```{r message=FALSE, echo=FALSE, warning=FALSE}
suppressPackageStartupMessages(library("ggplot2"))

# load results
path <- ("/Users/rmylonas/tmp/datamining_pumba/results/homodimers_1569571862.15343.RData")
load(path)

# load fasta file
fasta_seq_path <- "/Users/rmylonas/Work/PAF/projects/pumba/data/fasta/20170609_UP000005640_9606_with_headers.csv"
fasta_seqs <- read.csv(file=fasta_seq_path)

sample <- "HCT"
res <- results[[sample]]

show_mass_range <- c(10, 300)

peak_selection <- "closest_peak_masses" #"highest_peak_masses"
peak_selection_term <- "CLOSEST" # "HIGHEST"
```

### Data processing
- Data is extracted from Pumba
- Peaks are detected
- Data is annotated from additional sources (e.g. UniProt, SCOP)

### Peak detection
- Peaks below an absolute normalised intensity of 1e-08 are ignored.
- Only peaks >= 10% compared to strongest peak are retained.
- In between adjacent peaks the curve has to go through <= 80% of last peak, otherwise only the strongest peak is retained.  

## Proteins which migrated far away from theoretical mass
List of proteins with closest peak at more then 10kDa from theoretical mass.
Results are ordered by protein intensity.

```{r echo=FALSE, warning=FALSE}
  abs_peak_distances <- apply(res, 1, function(r){
    masses <- as.numeric(strsplit(as.character(r['peak_masses']), ",")[[1]])
    min(abs(masses - as.numeric(r['theo_weights'])))
  })

  flt <- res[abs_peak_distances > 10, ]
  flt <- flt[order(flt$prot_intensities, decreasing = TRUE),]
  
  # set the limits
  # summary(flt$theo_weights)
  # max(apply(flt, 1, function(r){  masses <- as.numeric(strsplit(as.character(r['peak_masses']), ",")[[1]]);  max(masses)}))
  # min(apply(flt, 1, function(r){  masses <- as.numeric(strsplit(as.character(r['peak_masses']), ",")[[1]]);  min(masses)}))
    # max(apply(flt, 1, function(r){  masses <- as.numeric(strsplit(as.character(r['peak_intensities']), ",")[[1]]);  max(masses)}))
  # min(apply(flt, 1, function(r){  masses <- as.numeric(strsplit(as.character(r['peak_intensities']), ",")[[1]]);  min(masses)}))
  
  
  mass_limits <- c(5.5, 710)
  int_limits <- c(1e-8, 3e-2)
  
  for(i in 1:nrow(flt)){
    l <- flt[i,]
    l$peak_masses
    masses <- as.numeric(strsplit(as.character(l$peak_masses), ",")[[1]])
    ints <- as.numeric(strsplit(as.character(l$peak_intensities), ",")[[1]])
    
    plot(1, type="n", xlab="", ylab="", xlim=mass_limits, ylim=int_limits, main=l$protein_acs, log="xy")
    # plot theo mass
    abline(v=l$theo_weights, col="blue")
    text(x=l$theo_weights, y=int_limits[1], col="blue", labels=round(l$theo_weights), adj=c(0.01,1))
    for(k in 1:length(ints)){
      segments(masses[k], 1e-8, masses[k], ints[k], col="red")
      points(x=masses[k], y=ints[k], col="red")
      text(x=masses[k], y=ints[k], col="red", labels=round(masses[k]), adj=c(0.01,0.01))
    }
    
    print(paste0("Protein AC: ", l$protein_acs))
    print(paste0("Fasta header:", fasta_seqs$header[as.character(fasta_seqs$ac) == as.character(l$protein_acs)]))
    print(paste0("Intensity: ", l$prot_intensities))
    print(paste0("Theo mol weight: ", l$theo_weights))
    print(paste0("PTMs: ", l$ptms))
    print(paste0("Signal pep: ", l$signal_peps))
    print(paste0("Glycosylations: ", l$glycosylations))
    print(paste0("Locations: ", l$locations))
    print(paste0("pIs: ", l$pIs))
    
  }

```



